Painting with Triangles

### 2014-04-22 paper.tex
foo

% uncomment one the following two lines, first one showing pictures
\documentclass[review]{acmsiggraph}
%\documentclass[review,draft]{acmsiggraph}

@@

% comment out this line to get figures; uncomment for fast latex
%\usepackage[draft]{graphicx}
\usepackage{graphicx}

@@

\subsection{Performance}
The current implementation uses Python and PyOpenGL. Python's performance is slower than
a lower level implementation in C++ would be. The drawing of the canvas is quick with
no lag. However, the actual process to take a stroke from rasterization to triangle
representation on the canvas can be slow. On a blank canvas, strokes take less than
one second to process. However, on a canvas with lots of geometry a stroke that overlaps
that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
necessary for the old points affected by the new stroke and the points associated with the
new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
calculations can take an arbitrarily long. This brings us to the final point, simplification.

\subsection{Simplification}
No matter how good the implementation, the fact that paintings can become arbitrarily complicated as more and more strokes are added
means there must be some way to simplify the mesh. In most situations where
many strokes overlap one another, most of the geometry is redundant. For example, when an opaque stroke is laid down on top of a complex geometry, that complex
geometry no longer has any useful information since the outline of the opaque stroke
completely defines it (all vertices are the same color).  Furthermore, even with translucent or feathered strokes, a large enough number of 
composited strokes may produce complex geometry that does not add much to the visual appearance of the canvas. Some of this geometry can be reduced without noticeable changes.  To accomplish this, our proposed simplification algorithm would identify a vertex that can be removed via an edge collapse by measuring the change the collapse would cause in the image.  The change is computed as the magnitude of the color shift and the translation of a boundary.  If these deltas are below a threshold, then the collapse can proceed.  This process could be done when a stroke is added to the canvas, or continually in a background thread to avoid increasing apparent latency.  Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue to iterate on a painting without the program becoming too slow.

@@

There are a number of intersting areas for future work, including:

%\noindent
%{\bf Simplification.}
%Our prototype implementation provides a baseline painting experience, in

Future work should focus on
implementing a simplification algorithm to allow the complexity of the drawings to grow. Other
improvements might focus on creating interesting manipulation tools, or other operations that utilize
the underlying triangle mesh structure to make the artists' job easier.

>>>>

% uncomment one the following two lines, first one showing pictures
%\documentclass[review]{acmsiggraph}
\documentclass[review,draft]{acmsiggraph}

@@

% comment out this line to get figures; uncomment for fast latex
\usepackage[draft]{graphicx}
%\usepackage{graphicx}

@@

\subsection{Performance}
The current implementation uses Python and PyOpenGL. Python's performance is slower than
a lower level implementation in C++ would be. The drawing of the canvas is quick with
no lag. However, the actual process to take a stroke from rasterization to triangle
representation on the canvas can be slow. On a blank canvas, strokes take less than
one second to process. However, on a canvas with lots of geometry a stroke that overlaps
that geometry can take several seconds to render. Some of this could be alleviated by a carefully optimized implementation. However, there are many calculations
necessary for the old points affected by the new stroke and the points associated with the
new stroke itself. Since these geometries can become arbitrarily complicated (in pathological cases), these
calculations can take an arbitrarily long.
% This brings us to the final point, simplification.

\subsection{Simplification}
\label{sec:simplify}
No matter how good the implementation, the fact that paintings can become arbitrarily complicated as more and more strokes are added
means there must be some way to simplify the mesh. In most situations where
many strokes overlap one another, most of the geometry is redundant. For example, when an opaque stroke is laid down on top of a complex geometry, that complex
geometry no longer has any useful information since the outline of the opaque stroke
completely defines it (all vertices are the same color).  Furthermore, even with translucent or feathered strokes, a large enough number of 
composited strokes may produce complex geometry that does not add much to the visual appearance of the canvas. Some of this geometry can be reduced without noticeable changes.  To accomplish this, our proposed simplification algorithm would identify a vertex that can be removed via an edge collapse by measuring the change the collapse would cause in the image.  The change is computed as the magnitude of the color shift and the translation of a boundary.  If these deltas are below a threshold, then the collapse can proceed.  This process could be done when a stroke is added to the canvas, or continually in a background thread to avoid increasing apparent latency.  Though we have yet to implement this simplification algorithm, it is important to ensure a painter can continue to iterate on a painting without the program becoming too slow.
%
Finally we note that Liao~\etal~\shortcite{10.1109/TVCG.2012.76} demonstrated that this kind of simplification can be effective at reducing mesh complexity, at least for natural imagery.
%
In their case they were working on an offline process for the entire image, whereas our
problem needs to be interactive on mouse-up (or as a background thread) but fortunately
only needs to operate on regions of the mesh that have changed during recent painting.

@@

Aside from the simplification challenge described in \sect{sec:simplify}, there are a number of other interesting areas for future work, including:

\noindent
{\bf Curved boundaries.}
In our implementation, curve boundaries are represented as polylines. We believe it would be relatively straightforward to modify these boundaries to be represented by higher-order curves, 
for example a Catmull-Rom spline passing through the vertices of the mesh. This would offer
higher-quality paths for the same sampling rate, but would require subdividing the path at
rendering time based on the image resolution. Today such operations can typically be 
performed on graphics hardware with little impact on runtime performance.

\noindent
{\bf Geometric operations.}
Because the underlying representation is vector in nature, it should robustly support warping and
other geometric operations. It should also be easy to provide other kinds of useful tools to the 
artist like a ``smoothing brush'' which could locally adjust the geometry and/or the colors to soften a region. Finally, it would be interesting to offer this kind of painting interface on non-planar canvasses, for example allowing an artist to paint directly onto a manifold representing a shape in 3D.
